---
title: "Practical Machine Learning: Course Assignment"
output: html_document
---

#SYNOPSIS

The purpose of this assignment is to build a prediction model on the weight lifting data set. There is a training and test data set available in the links below.  The variable we are attempting to predict is "classe" (null in the test set).

##LOAD THE DATA

The data is loaded as follows:

```{r download_files, cache=TRUE, warning=FALSE, message=FALSE}

library(caret)
library(kernlab)
library(ElemStatLearn)

dat_train <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")

dat_test <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")

dim(dat_train)
dim(dat_test)

```

##IDENTIFY PREDICTORS

According to the assignment instructions, we are using "data from the accelerometers on the belt, forearm, arm, and dumbell of 6 participants".  The applicable variables in the training data set will includes these terms, so we modify the data sets to include only those variables plus "classe". Additionally we are only concerned with the columns in the test set which have values and will reduce the test and training datasets.


```{r modify_data, cache=TRUE, warning=FALSE, message=FALSE}


#dat_test <- dat_test[,grep("arm|dumbbell|belt|classe",colnames(dat_test))]
#dat_test<- dat_test[,grep("accel|classe",colnames(dat_test))]
dat_test <- dat_test[,colSums(is.na(dat_test)) != nrow(dat_test)]
dat_test$classe <- ""
dat_test <- dat_test[,c(grep("arm|dumbbell|belt|classe",colnames(dat_test)))]


dat_train <- dat_train[, colnames(dat_test)]
#dat_train <- dat_train[,grep(names,colnames(dat_train))]
#dat_train <- dat_train[,grep("accel|classe",colnames(dat_train))]



```

##BUILD AND TEST A PREDICTION MODEL

Because the training set contains only valid entries, we will partition this into a training and test test - this is a cross validation process.  

We will predict the outcome of "classe" using all predictos with a boosting model, then validate its accuracy against the testing data set.


```{r build, cache=TRUE, warning=FALSE, message=FALSE, include=FALSE}

inTrain<- createDataPartition(dat_train$classe, p=.5, list=FALSE)
training <- dat_train[inTrain,]
testing <- dat_train[-inTrain,]


modFitGBM <- train(classe ~ ., data=training, method="gbm")
predictGBM <- predict(modFitGBM, testing)


```


By compiling a confusion matrix, we can estimate the accuracy of the prediction model vs. the actual testing output.  This is a reasonable estimate of the out of sample error rate.  The variable importance (varImp()) provides a guideline as to which variables have the largest impact on predicting outcomes.


```{r outofsample, cache=TRUE, warning=FALSE, message=FALSE}

confusionMatrix(testing$classe, predictGBM)$overall[1]
varImp(modFitGBM)

```


#RESULTS

The results of the prediction model:

```{r results, cache=TRUE, warning=FALSE, message=FALSE}

predictDAT_TEST <- predict(modFitGBM, dat_test)
predictDAT_TEST

```




